Optimal (Multiway) Spatial Joins

### 2024-05-29 4.2.tex

\noindent \underline{\em Proof of Statement (2).} 
First, we deviate a little to prove the following statement. For any rectangle $r\in \bigcup_{i = 1}^{k-2}$, if a segment $h\in H$ crosses $r$, then $h$ also crosses the trimmed rectangle of $r$, denoted as $r'$. Since the x-range of $r'$ is a subset of the x-range of $r$, it suffices to show that the y-coordinate of $h$ is in the range $[\ybot(r'), \ytop(r')]$. Let $h = [x_1,x_2]\times y$. Since $\ybot(r')$ is the y-coordinate of the lowest segment in $H$ that crosses $r$, and $h$ crosses $r$, we have $\ybot(r') \leq y$. Moreover, since $\ytop(r')$ is the y-coordinate of the highest segment in $H$ that crosses $r$, and $h$ crosses $r$, we have $y \leq \ytop(r')$. Therefore, $y \in [\ybot(r'), \ytop(r')]$ and $h$ crosses $r'$.  Analogously, we can prove that if a segment $v\in V$ crosses $r$, then $v$ also crosses the trimmed rectangle of $r$.

@@

Now, we prove Statement (2). Since $(r_1,...,r_{k-2},h,v)\in \J_1$, both segments $h$ and $v$ cross all of the rectangles $r_1,..., r_{k-2}$. By the discussion above, $h$ and $v$ also cross the trimmed rectangles $r_1',..., r_{k-2}'$. Therefore, the point $h \cap v$ is in $r_i'$, for each $i \in [k-2]$. This implies that $\bigcap_{i = 1}^{k-2}r_i' \neq \emptyset$, and $(r_1',..., r_{k-2}')\in \J(R_1',...,R_{k-2}')$.

% The fact $(r_1,...,r_{k-2},h,v)\in \J_1$ indicates that both segments $h$ and $v$ cross all of the rectangles $r_1,..., r_{k-2}$ which means that for $i\in[k-2]$, the y-coorninate of $h$ is in the range $[\ybot(r_i'), \ytop(r_i')]$ and the x-coordinate of $v$ is in $[\xleft(r_i'), \xright(r_i')]$. Additionally, for each $i\in[k-2]$, the x-range (resp. y-range) of $r_i'$ is a subset of the x-range (resp. y-range) of $r_i$, so $h$ (resp. $v$) crosses the rectangles $r_i'$. Hence, the point $h \cap v$ is in $B_\bm{t} = \bigcap_{i = 1}^{k-2}r_i'$, and $h, v$ cross $B_{\bm{t}}$. The former implies that $\bm{t} \in \J(R_1',...,R_{k-2}')$, and the latter implies that $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.

@@

\noindent \underline{\em Proof of Statement (3).}
For each $\bm{t'}\in  \J(R_1',...,R_{k-2}')$ and each $h \in \dcross_H(\bm{t'})$, we will identify a unique tuple $\bm{t}$ in $\J(R_1,...,R_{k-2},H,V)$ defined as follows. Let $r_i$ be the full rectangle of $r_i'$ for $i \in [k]$, and $v$ be the segment $v_\vdash$ of $\gleft(\bm{t'})$. Let $\bm{t} = (r_1,...,r_{k-2},h,v)$. Since $h$ crosses $B_{\bm{t'}}$ by the definition of $\dcross_H(\bm{t'})$, the $h\cap v$ is a point in $B_{\bm{t'}}$.
Therefore, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$. he result tuple is unique because for any two distinct tuples $\bm{t'_1}$ and $\bm{t'_2}$ in $\J(R_1',...,R_{k-2}')$, there exists an integer $i \in [k-2]$ s.t. the full rectangles of $\bm{t'_1}[i]$ and $\bm{t'_2}[i]$ are different, which means the $i$-th components of their associating tuples $\bm{t_1}, \bm{t_2} \in \J(R_1,...,R_{k-2},H,V)$ are different. As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_H(\bm{t})| \leq \out$.

Similarly, for each $\bm{t'}\in  \J(R_1',...,R_{k-2}')$ and each $v \in \dcross_V(\bm{t'})$, we can identify a unique tuple $\bm{t}$ in $\J(R_1,...,R_{k-2},H,V)$ defined as follows. Let $r_i$ be the full rectangle of $r_i'$ for $i \in [k]$, and $h$ be the segment $h_\bot$ of $\gbot(\bm{t'})$. Let $\bm{t} = (r_1,...,r_{k-2},h,v)$. Since $v$ crosses $B_{\bm{t'}}$, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$. As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_V(\bm{t})| \leq \out$.

>>>>

@@

\noindent \underline{\em Proof of Statement (3).} Consider an arbitrary $\bm{t} \in \J(R_1',...,R_{k-2}')$. First, we claim that $|\dcross_H(\bm{t})| \neq \emptyset$. Let $r' = \gbot(\bm{t})$, and let $r$ be the full rectangle of $r'$. Let $h$ be the lowest segment that crosses $r$. We will prove the claim by showing that $h \in \dcross_H(\bm{t})$. More specifically, we will prove that $h$ crosses $B_{\bm{t}}$ and $r'$. From the first paragraph of the proof of Statement (2), we know that $h$ crosses $r'$ because $h$ crosses $r$.
By how $r'$ is defined, the bottom edge of $r'$ is contained in $h$. Since $r' = \gbot(\bm{t})$, the bottom edge of $B_\bm{t}$ is contained in the bottom edge of $r'$, and hence in $h$. Thus, $h \cap B_{\bm{t}}\neq \emptyset$. Since $h$ crosses $r'$, $r'$ does not contain any of the endpoints of $h$. As $r'$ covers $B_{\bm{t}}$, $B_{\bm{t}}$ does not contain any of the endpoints of $h$. Thus, $h$ crosses $B_{\bm{t}}$. This proves that $h \in \dcross_H(\bm{t})$, and hence $|\dcross_H(\bm{t})| \neq \emptyset$. By symmetry, $|\dcross_V(\bm{t})|$ is also nonempty. Equivalently, we have $|\dcross_H(\bm{t})| \geq 1$ and $|\dcross_V(\bm{t})| \geq 1$.

@@

Therefore, to prove Statement (3),
it suffices to show that \\
\myeqn{
    \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}(|\dcross_H(\bm{t})|\cdot |\dcross_V(\bm{t})|) \leq \out. \label{eqn:lmm4.2.3}
    } Recall that $\J^*$ is the set of tuples generated by the algorithm \textbf{generate-$\J^*$}(see below Example 4.3). By the discussion following the algorithm, we have $\J^*\subseteq \J(R_1,...,R_{k-2},H,V)$, which means $|\J^*|\leq \out$. Thus, we can prove Inequality~\eqref{eqn:lmm4.2.3} by showing that its left-hand side is equal to $|\J^*|$.
 By the line 2-4 of \textbf{generate-$\J^*$}, the number of tuples generated in line 5 is exactly the left-hand side of Inequality~\eqref{eqn:lmm4.2.3}.
Besides, it is obvious that every tuple $(r_1,...,r_{k-2},h,v)$ generated in line 5 of \textbf{generate-$\J^*$} is unique. Therefore, Inequality~\eqref{eqn:lmm4.2.3} holds. This completes the proof of Statement (3).
